---
title: "p8105_hw1_rl3411.Rmd"
author: "rl3411"
date: "2023-09-14"
output: github_document
---

```{r load package and data, message = FALSE}
# load all packages required and setup dataset
library(moderndive) 
library(ggplot2)
library(tidyverse)
data("early_january_weather")
```

# Problem 1

## About the dataset

The `early_january_weather` dataset includes hourly meteorological data for EWR, Newark Liberty International Airport, for the month of January 2013.  This dataset has `r ncol(early_january_weather)` variables (thus `r ncol(early_january_weather)` columns) and `r nrow(early_january_weather)` rows of data. The variables given are `r colnames(early_january_weather)`, and the description of each variable are listed in the table below.

The mean temperature is `r round(mean(early_january_weather$temp),digits=1)` degrees, and ranges from `r round(min(early_january_weather$temp),digits=1)`degrees to `r round(max(early_january_weather$temp),digits=1)`degrees, all to the nearest decimal. This dataset is mostly complete, except `wind_dir` with 5 NAs, `wind_gust` with 308 NAs, `pressure` with 38 NAs.


| Variable name in dataset                | Description                                                |
| --------------------------------------- | ---------------------------------------------------------- |
| `origin`                                | Weather station                                            |
| `year`, `month`, `day`, `hour`          | Time of recording                                          |
| `temp`, `dewp`                          | Temperature and dewpoint in F                              |
| `humid`                                 | Relative humidity                                          |
| `wind_dir`, `wind_speed`, `wind_gust`   | Wind direction (in degrees), speed and gust speed (in mph) |
| `precip`                                | Precipitation, in inches                                   |
| `pressure`                              | Sea level pressure in millibars                            |
| `visib`                                 | Visibility in miles                                        |
| `time_hour`                             | Date and hour of the recording                             |
  
## Scatterplot: `temp` vs `time_hour`

Looking at the plot below, we see that the temperature was fluctuating between hours, but showed a general positive trend through time, indicating that it was getting warmer. The humidity was very high between January 11 to January 14, suggesting a rainy period. After that, the temperature peaked at almost 60 degrees before taking a dip down to approximately 37 degrees.
```{r scatterplot, message = FALSE}
weather_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp)) + 
  geom_point(aes(colour = humid))
weather_plot
ggsave("weather_plot.png")
```

# Problem 2

## Create dataframe

Let us first create a dataframe.

```{r dataframe}
prob2 <- tibble(
  rand_samp = rnorm(10, 0, 1),
  logical_vec = rand_samp > 0,
  char_vec = c('are','we','supposed','to','give','an','example','i','dont','know'),
  factor_vec = factor(c('yes','no','maybe','no','no','maybe','yes','no','yes','yes'))
)
```

## Calculating means

Now, we shall try to take the mean of each variable:

1. We see that this works for `rand_samp`.
```{r mean1}
mean(pull(prob2, rand_samp))
```

2. This also works for `logical_vec`.
```{r mean2}
mean(pull(prob2, logical_vec))
```

3. However, taking the mean of the character variable `char_vec` returns a warning. This fails because we can only take means of numeric or logical variables, which `char_vec` is a character variable.
```{r mean3}
mean(pull(prob2, char_vec))
```

4. Lastly, taking the mean of `factor_vec` gives the same warning as the previous character variable, as this is a factor variable.
```{r mean4}
mean(pull(prob2, factor_vec))
```

## Conversion to numeric variables

```{r convert, message = FALSE}
as.numeric(pull(prob2, logical_vec))
as.numeric(pull(prob2, char_vec))
as.numeric(pull(prob2, factor_vec))
```

By applying `as.numeric` to those 3 variables, this converts `logical_vec` to a series of binary numbers 0 or 1. 1 corresponds to true and 0 corresponds to false. For the factor vector, it changes the levels into numbers 1, 2 and 3. R automatically arranges the levels alphabetically, therefore "**m**aybe" takes value 1, "**n**o" takes value 2, and "**y**es" takes value 3. Lastly, we cannot convert the character vector into numeric therefore it returns a series a NAs. We can first make this into a factor then convert it to a numeric vector by executing `as.numeric(factor(pull(prob2, char_vec)))`, but this just assigns a number to the alphabetical arrangement of the elements so taking the mean would not necessarily be useful. In general, we cannot make calculations on non-numeric variables, thus `as.numeric()` allows us to first convert variables into numeric then execute any necessary calculations.
